---
title: "NFL_Elo_Analysis"
output: html_document
date: "2022-10-30"
---
dataset:
```{r}
nfl_elo <- nfl_elo_proj

nfl_elo
```

creating final_score variable
```{r}
nfl_elo <- nfl_elo %>% 
  mutate(final_score = score_home_team - score_away_team)

nfl_elo
```

creating a division variable
```{r}
divisionPlacement = function(teamName){
  if(teamName == "BUF" | teamName == "NYJ" | teamName == "MIA" | teamName == "NE"){
    return("AFC EAST")
  }else if(teamName == "BAL" | teamName == "CIN" | teamName == "CLE" | teamName == "PIT"){
    return("AFC NORTH")
  }else if(teamName == "TEN" | teamName == "IND" | teamName == "JAX" | teamName == "HOU"){
    return("AFC SOUTH")
  }else if(teamName == "KC" | teamName == "LAC" | teamName == "DEN" | teamName == "OAK"){
    return("AFC WEST")
  }else if(teamName == "PHI" | teamName == "NYG" | teamName == "DAL" | teamName == "WSH"){
    return("NFC EAST")
  }else if(teamName == "GB" | teamName == "MIN" | teamName == "DET" | teamName == "CHI"){
    return("NFC NORTH")
  }else if(teamName == "ATL" | teamName == "CAR" | teamName == "TB" | teamName == "NO"){
    return("NFC SOUTH")
  }else if(teamName == "SF" | teamName == "SEA" | teamName == "LAR" | teamName == "ARI"){
    return("NFC WEST")
  }else{
    return("ERROR")
  }
}
nfl_elo %>% 
  group_by(home_team) %>%
  mutate(home_team_division = divisionPlacement(home_team[1])) %>% 
  group_by(away_team) %>% 
  mutate(away_team_division = divisionPlacement(away_team[1])) -> division_df
division_df
```

creating total games played variable
```{r}
nfl_elo %>% 
  group_by(home_team) %>% 
  summarize(homeGames=n()) -> homeGames

nfl_elo %>% 
  group_by(away_team) %>% 
  summarize(awayGames=n()) -> awayGames

merge(awayGames, homeGames, by.x = "away_team", by.y = "home_team") %>% 
  group_by(away_team) %>% 
  summarize(totalGames = awayGames + homeGames) %>% 
  rename_at("away_team", ~"team") 
```

Which team had the highest avg pts per game
```{r}
nfl_elo %>% 
  group_by(home_team) %>% 
  summarise(total_home_points = sum(score_home_team)) -> total_home_points

nfl_elo %>% 
  group_by(away_team) %>% 
  summarise(total_away_points = sum(score_away_team)) -> total_away_points

totalPoints <- merge(total_home_points, total_away_points, by.x = "home_team", by.y = "away_team") %>% 
  group_by(home_team) %>% 
  summarize(totalPoints = total_home_points + total_away_points) %>% 
  rename_at("home_team", ~"team")

nfl_elo %>% 
  group_by(home_team) %>% 
  summarize(homeGames=n()) -> homeGames

nfl_elo %>% 
  group_by(away_team) %>% 
  summarize(awayGames=n()) -> awayGames

totalGames <- merge(awayGames, homeGames, by.x = "away_team", by.y = "home_team") %>% 
  group_by(away_team) %>% 
  summarize(totalGames = awayGames + homeGames) %>% 
  rename_at("away_team", ~"team")

left_join(totalPoints, totalGames, by = "team") %>% 
  mutate(avg_points_per_game = totalPoints / totalGames) %>% 
  select(team, avg_points_per_game) %>% 
  arrange(-avg_points_per_game) %>% 
  mutate_if(is.numeric, round)
```

Which team had the highest avg elo per game
```{r}
nfl_elo %>% 
  group_by(home_team) %>% 
  summarise(total_home_elo = sum(elo_home_team)) -> total_home_elo

nfl_elo %>% 
  group_by(away_team) %>% 
  summarise(total_away_elo = sum(elo_away_team)) -> total_away_elo

totalelo <- merge(total_home_elo, total_away_elo, by.x = "home_team", by.y = "away_team") %>% 
  group_by(home_team) %>% 
  summarize(totalelo = total_home_elo + total_away_elo) %>% 
  rename_at("home_team", ~"team")

nfl_elo %>% 
  group_by(home_team) %>% 
  summarize(homeGames=n()) -> homeGames

nfl_elo %>% 
  group_by(away_team) %>% 
  summarize(awayGames=n()) -> awayGames

totalGames <- merge(awayGames, homeGames, by.x = "away_team", by.y = "home_team") %>% 
  group_by(away_team) %>% 
  summarize(totalGames = awayGames + homeGames) %>% 
  rename_at("away_team", ~"team")

left_join(totalelo, totalGames, by = "team") %>% 
  mutate(avg_elo_per_game = totalelo / totalGames) %>% 
  select(team, avg_elo_per_game) %>% 
  arrange(-avg_elo_per_game) %>% 
  mutate_if(is.numeric, round)
```

which qb had the highest avg points per game
```{r}
nfl_elo %>% 
  group_by(home_qb) %>% 
  summarise(total_home_qb_points = sum(score_home_team)) -> total_home_qb_points

nfl_elo %>% 
  group_by(away_qb) %>% 
  summarise(total_away_qb_points = sum(score_away_team)) -> total_away_qb_points

totalQBPoints <- merge(total_home_qb_points, total_away_qb_points, by.x = "home_qb", by.y = "away_qb") %>% 
  group_by(home_qb) %>% 
  summarize(totalQBPoints = total_home_qb_points + total_away_qb_points) %>% 
  rename_at("home_qb", ~"QB")

nfl_elo %>% 
  group_by(home_qb) %>% 
  summarize(homeQBGames=n()) -> homeQBGames

nfl_elo %>% 
  group_by(away_qb) %>% 
  summarize(awayQBGames=n()) -> awayQBGames

totalQBGames <- merge(awayQBGames, homeQBGames, by.x = "away_qb", by.y = "home_qb") %>% 
  group_by(away_qb) %>% 
  summarize(totalQBGames = awayQBGames + homeQBGames) %>% 
  rename_at("away_qb", ~"QB") 

left_join(totalQBGames, totalQBPoints, by = "QB") %>% 
  filter(totalQBGames > 16) %>% 
  mutate(avg_qb_points_per_game = totalQBPoints / totalQBGames) %>% 
  select(QB, avg_qb_points_per_game) %>% 
  arrange(-avg_qb_points_per_game) %>% 
  mutate_if(is.numeric, round)
```

which qb had the highest elo per game
```{r}
nfl_elo %>% 
  group_by(home_qb) %>% 
  summarise(total_home_qb_elo = sum(qb1_value_post)) -> total_home_qb_elo

nfl_elo %>% 
  group_by(away_qb) %>% 
  summarise(total_away_qb_elo = sum(qb2_value_post)) -> total_away_qb_elo

totalQBelo <- merge(total_home_qb_elo, total_away_qb_elo, by.x = "home_qb", by.y = "away_qb") %>% 
  group_by(home_qb) %>% 
  summarize(totalQBelo = total_home_qb_elo + total_away_qb_elo) %>% 
  rename_at("home_qb", ~"QB")

nfl_elo %>% 
  group_by(home_qb) %>% 
  summarize(homeQBGames=n()) -> homeQBGames

nfl_elo %>% 
  group_by(away_qb) %>% 
  summarize(awayQBGames=n()) -> awayQBGames

totalQBGames <- merge(awayQBGames, homeQBGames, by.x = "away_qb", by.y = "home_qb") %>% 
  group_by(away_qb) %>% 
  summarize(totalQBGames = awayQBGames + homeQBGames) %>% 
  rename_at("away_qb", ~"QB") 

left_join(totalQBGames, totalQBelo, by = "QB") %>% 
  filter(totalQBGames > 16) %>% 
  mutate(avg_qb_elo_per_game = totalQBelo / totalQBGames) %>% 
  select(QB, avg_qb_elo_per_game) %>% 
  arrange(-avg_qb_elo_per_game) %>% 
  mutate_if(is.numeric, round)
```

which division has the highest avg team elo
```{r}
division_df %>% 
  group_by(home_team_division) %>% 
  summarise(total_h_division_elo = sum(elo_home_team)) -> total_h_division_elo

division_df %>% 
  group_by(away_team_division) %>% 
  summarise(total_a_division_elo = sum(elo_away_team)) -> total_a_division_elo

totalelo <- merge(total_h_division_elo, total_a_division_elo, by.x = "home_team_division", by.y = "away_team_division") %>% 
  group_by(home_team_division) %>% 
  summarize(totalelo = total_h_division_elo + total_a_division_elo) %>% 
  rename_at("home_team_division", ~"division")

division_df %>% 
  group_by(home_team_division) %>% 
  summarize(homeGames=n()) -> homeGames

division_df %>% 
  group_by(away_team_division) %>% 
  summarize(awayGames=n()) -> awayGames

totalGames <- merge(awayGames, homeGames, by.x = "away_team_division", by.y = "home_team_division") %>% 
  group_by(away_team_division) %>% 
  summarize(totalGames = awayGames + homeGames) %>% 
  rename_at("away_team_division", ~"division")

left_join(totalelo, totalGames, by = "division") %>% 
  mutate(avg_elo_per_game = totalelo / totalGames) %>% 
  select(division, avg_elo_per_game) %>% 
  arrange(-avg_elo_per_game) %>% 
  mutate_if(is.numeric, round) -> div_elo

div_elo
```

Total team wins 
```{r}
nfl_elo %>% 
  mutate(finalScore = score_home_team - score_away_team) -> finalScore

finalScore %>% 
  filter(finalScore>0) %>% 
  group_by(home_team) %>% 
  summarize(homeTeamWins = n()) -> homeTeamWins

finalScore %>% 
  filter(finalScore < 0) %>% 
  group_by(away_team) %>% 
  summarize(awayTeamWins = n()) -> awayTeamWins

merge(awayTeamWins, homeTeamWins, by.x = "away_team", by.y = "home_team") %>% 
  group_by(away_team) %>% 
  summarize(totalWins = awayTeamWins + homeTeamWins) %>% 
  rename_at("away_team", ~"team") -> totalWins

totalWins %>% 
  arrange(desc(totalWins))
  
```

Total games each team has played
```{r}
nfl_elo %>% 
  group_by(home_team) %>% 
  summarize(homeGames=n()) -> homeGames

nfl_elo %>% 
  group_by(away_team) %>% 
  summarize(awayGames=n()) -> awayGames

merge(awayGames, homeGames, by.x = "away_team", by.y = "home_team") %>% 
  group_by(away_team) %>% 
  summarize(totalGames = awayGames + homeGames) %>% 
  rename_at("away_team", ~"team") -> totalGames
```

Win percentage by team
```{r}
left_join(totalGames, totalWins, by = "team") %>% 
  mutate(win_percent = totalWins / totalGames) -> win_percentage

win_percentage
```

bar of top teams win%
```{r}
win_percentage %>% 
  arrange(-win_percent) %>% 
  head(10) %>% 
  ggplot(mapping = aes(x = team, y = win_percent, fill = team)) +
  geom_col() +
  theme_light() +
  xlab("Team") + ylab("Win %") +
  ggtitle("Win % for the 10 Most Winning Teams") 
```

bar chart of divisions avg elo
```{r}
div_elo %>% 
  ggplot(mapping = aes(x = division, y = avg_elo_per_game, fill = division)) +
  geom_col() +
  theme_light() +
  coord_cartesian(ylim = c(1400,1600)) + 
  ggtitle("Average Elo by Division") +
  xlab("Division") + ylab("Average Elo per Game") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

which division has the highest record per year
```{r}
nfl_elo %>% 
  mutate(finalScore = score_home_team - score_away_team) -> finalScore

finalScore %>% 
  filter(finalScore>0) %>% 
  group_by(home_team, season) %>% 
  summarize(homeTeamWins = n()) -> homeTeamWins

finalScore %>% 
  filter(finalScore < 0) %>% 
  group_by(away_team, season) %>%  
  summarize(awayTeamWins = n()) -> awayTeamWins

merge(awayTeamWins, homeTeamWins, by.x = c("away_team", "season"), by.y = c("home_team", "season")) %>%
  group_by(away_team, season) %>% 
  summarize(totalWins = awayTeamWins + homeTeamWins) %>% 
  rename_at("away_team", ~"team") -> seasonTeamWins

seasonTeamWins %>% 
  arrange(season, desc(totalWins)) %>%
  mutate(division = divisionPlacement(team[1])) %>% 
  arrange(season, division, desc(totalWins)) -> divSeasonTeamWins

divSeasonTeamWins %>% 
  group_by(season, division) %>% 
  slice_max(totalWins, with_ties = FALSE) %>% 
  arrange(division) -> divwinnerbyyear

divwinnerbyyear
```

bar chart of total wins of each division winner
```{r}
divwinnerbyyear %>% 
  ggplot(mapping = aes(x = totalWins, y = division)) +
  geom_col(mapping = aes(fill = division)) +
  coord_cartesian(xlim = c(0, 20)) + 
  facet_wrap(~ season) +
  theme(axis.text = element_text(size = 6))
```

which division has the highest avg team points
```{r}
division_df %>% 
  group_by(home_team_division) %>% 
  summarise(total_h_division_pts = sum(score_home_team)) -> total_h_division_pts

division_df %>% 
  group_by(away_team_division) %>% 
  summarise(total_a_division_pts = sum(score_away_team)) -> total_a_division_pts

totalpts <- merge(total_h_division_pts, total_a_division_pts, by.x = "home_team_division", by.y = "away_team_division") %>% 
  group_by(home_team_division) %>% 
  summarize(totalpts = total_h_division_pts + total_a_division_pts) %>% 
  rename_at("home_team_division", ~"division")

division_df %>% 
  group_by(home_team_division) %>% 
  summarize(homeGames=n()) -> homeGames

division_df %>% 
  group_by(away_team_division) %>% 
  summarize(awayGames=n()) -> awayGames

totalGames <- merge(awayGames, homeGames, by.x = "away_team_division", by.y = "home_team_division") %>%
  group_by(away_team_division) %>% 
  summarize(totalGames = awayGames + homeGames) %>% 
  rename_at("away_team_division", ~"division")

left_join(totalpts, totalGames, by = "division") %>% 
  mutate(avg_pts_per_game = totalpts / totalGames) %>% 
  select(division, avg_pts_per_game) %>% 
  arrange(-avg_pts_per_game) 
```

which division has the highest avg points per game
```{r}
division_df %>% 
  group_by(home_team_division) %>% 
  summarise(total_h_division_points = sum(score_home_team)) -> total_h_division_points

division_df %>% 
  group_by(away_team_division) %>% 
  summarise(total_a_division_points = sum(score_away_team)) -> total_a_division_points

totalPoints <- merge(total_h_division_points, total_a_division_points, by.x = "home_team_division", by.y = "away_team_division") %>% 
  group_by(home_team_division) %>% 
  summarize(totalPoints = total_h_division_points + total_a_division_points) %>% 
  rename_at("home_team_division", ~"division")

division_df %>% 
  group_by(home_team_division) %>% 
  summarize(homeGames=n()) -> homeGames

division_df %>% 
  group_by(away_team_division) %>% 
  summarize(awayGames=n()) -> awayGames

totalGames <- merge(awayGames, homeGames, by.x = "away_team_division", by.y = "home_team_division") %>% 
  group_by(away_team_division) %>% 
  summarize(totalGames = awayGames + homeGames) %>% 
  rename_at("away_team_division", ~"division")

left_join(totalPoints, totalGames, by = "division") %>% 
  mutate(avg_points_per_game = totalPoints / totalGames) %>% 
  select(division, avg_points_per_game) %>% 
  arrange(-avg_points_per_game)
```

which division has the best qb play by avg qb elo per game
```{r}
division_df %>% 
  group_by(home_team_division) %>% 
  summarise(total_home_div_elo = sum(qb1_value_post)) -> total_home_div_elo

division_df %>% 
  group_by(away_team_division) %>% 
  summarise(total_away_div_elo = sum(qb2_value_post)) -> total_away_div_elo

totalQBelo <- merge(total_home_div_elo, total_away_div_elo, by.x = "home_team_division", by.y = "away_team_division") %>% 
  group_by(home_team_division) %>% 
  summarize(totalQBelo = total_home_div_elo + total_away_div_elo) %>% 
  rename_at("home_team_division", ~"division")

division_df %>% 
  group_by(home_team_division) %>% 
  summarize(homeQBGames=n()) -> homeQBGames

division_df %>% 
  group_by(away_team_division) %>% 
  summarize(awayQBGames=n()) -> awayQBGames

totalQBGames <- merge(awayQBGames, homeQBGames, by.x = "away_team_division", by.y = "home_team_division") %>% 
  group_by(away_team_division) %>% 
  summarize(totalQBGames = awayQBGames + homeQBGames) %>% 
  rename_at("away_team_division", ~"division") 

left_join(totalQBGames, totalQBelo, by = "division") %>% 
  filter(totalQBGames > 16) %>% 
  mutate(avg_qb_elo_per_game = totalQBelo / totalQBGames) %>% 
  select(division, avg_qb_elo_per_game) %>% 
  arrange(-avg_qb_elo_per_game)
```

best team by division by elo
```{r}
merge(totalQBelo, totalQBGames) %>% 
  group_by(QB) %>% 
  mutate(avgQBelo = (totalQBelo/totalQBGames)) %>% 
  arrange(desc(avgQBelo)) %>% 
  group_by(QB) %>%
  mutate(division = divisionPlacement(team[1])) -> QBeloByDivision

QBeloByDivision %>% 
  group_by(division) %>% 
  filter(avgQBelo == max(avgQBelo))

```
```{r}
distinct_QB <- merge(total_home_qb_elo, total_away_qb_elo, by.x = "home_qb", by.y = "away_qb") %>% 
  group_by(home_qb) %>% 
  mutate(totalQBelo = total_home_qb_elo + total_away_qb_elo) %>% 
  rename_at("home_qb", ~"QB") %>% 
  distinct(QB, season.x)
```


```{r}
totalelo
totalGames
```

```{r}
division_df %>% 
  group_by(home_qb) %>% 
  mutate(total_home_qb_elo = sum(qb1_value_post)) -> total_home_qb_elo

division_df %>% 
  group_by(away_qb) %>% 
  mutate(total_away_qb_elo = sum(qb2_value_post)) -> total_away_qb_elo

totalQBelo <- merge(total_home_qb_elo, total_away_qb_elo, by.x = "home_qb", by.y = "away_qb") %>% 
  group_by(home_qb) %>% 
  mutate(totalelo = total_home_qb_elo + total_away_qb_elo) %>% 
  rename_at("home_qb", ~"QB") %>% 
  rename_at("home_team.x", ~"team") %>% 
  rename_at("season.x", ~"season") %>% 
  rename_at("home_team_division.x", ~"division") %>% 
  select(QB, season, team, division, totalelo) %>% 
  distinct(QB, season, .keep_all = TRUE)


division_df %>% 
  group_by(home_qb) %>% 
  mutate(homeQBGames=n()) -> homeQBGames

division_df %>% 
  group_by(away_qb) %>% 
  mutate(awayQBGames=n()) -> awayQBGames

totalQBGames <- merge(awayQBGames, homeQBGames, by.x = "away_qb", by.y = "home_qb") %>% 
  group_by(away_qb, season.x) %>% 
  mutate(totalGames = awayQBGames + homeQBGames) %>% 
  rename_at("away_qb", ~"QB") %>% 
  rename_at("home_team.x", ~"team") %>% 
  rename_at("season.x", ~"season") %>% 
  rename_at("home_team_division.x", ~"division") %>% 
  select(QB, season, team, division, totalGames) %>% 
  distinct(QB, .keep_all = TRUE)


left_join(totalQBGames, totalQBelo, by = c("QB", "season")) %>% 
  filter(totalGames > 16) %>% 
  mutate(avg_qb_elo_per_game = totalelo / totalGames) 
```

highest avg points per game by division 
```{r}
division_df %>% 
  group_by(home_team) %>% 
  mutate(total_home_elo = sum(elo_home_team)) -> total_home_elo

division_df %>% 
  group_by(away_team) %>% 
  mutate(total_away_elo = sum(elo_away_team)) -> total_away_elo

totalelo <- merge(total_home_elo, total_away_elo, by.x = "home_team", by.y = "away_team") %>% 
  rename_at("home_team", ~"team") %>%  
  rename_at("home_team_division.x", ~"division") %>% 
  group_by(team) %>% 
  mutate(totalelo = total_home_elo + total_away_elo) %>% 
  select(team, division, totalelo) %>% 
  distinct(team, .keep_all = TRUE)

division_df %>% 
  group_by(home_team) %>% 
  mutate(homeGames=n()) -> homeGames

division_df %>% 
  group_by(away_team) %>% 
  mutate(awayGames=n()) -> awayGames

totalGames <- merge(awayGames, homeGames, by.x = "away_team", by.y = "home_team") %>% 
  rename_at("away_team", ~"team") %>% 
  rename_at("away_team_division.x", ~"division") %>% 
  group_by(team) %>% 
  mutate(totalGames = awayGames + homeGames) %>% 
  select(team, division, totalGames) %>% 
  distinct(team, .keep_all = TRUE)

left_join(totalelo, totalGames, by = "team") %>% 
  mutate(avg_elo_per_game = totalelo / totalGames) %>% 
  rename_at("division.x", ~"division") %>% 
  select(-division.y, -totalelo, -totalGames) %>% 
  group_by(division) %>% 
  slice_max(avg_elo_per_game)  %>% 
  mutate_if(is.numeric, round)
```

highest avg points per game by division 
```{r}
division_df %>% 
  group_by(home_team) %>% 
  mutate(total_home_pts = sum(score_home_team)) -> total_home_pts

division_df %>% 
  group_by(away_team) %>% 
  mutate(total_away_pts = sum(score_away_team)) -> total_away_pts

totalpts <- merge(total_home_pts, total_away_pts, by.x = "home_team", by.y = "away_team") %>% 
  rename_at("home_team", ~"team") %>% 
  rename_at("home_team_division.x", ~"division") %>% 
  group_by(team) %>% 
  mutate(totalpts = total_home_pts + total_away_pts) %>% 
  select(team, division, totalpts) %>% 
  distinct(team, .keep_all = TRUE)

division_df %>% 
  group_by(home_team) %>% 
  mutate(homeGames=n()) -> homeGames

division_df %>% 
  group_by(away_team) %>% 
  mutate(awayGames=n()) -> awayGames

totalGames <- merge(awayGames, homeGames, by.x = "away_team", by.y = "home_team") %>% 
  rename_at("away_team", ~"team") %>% 
  rename_at("away_team_division.x", ~"division") %>% 
  group_by(team) %>% 
  mutate(totalGames = awayGames + homeGames) %>% 
  select(team, division, totalGames) %>% 
  distinct(team, .keep_all = TRUE)

left_join(totalpts, totalGames, by = "team") %>% 
  mutate(avg_pts_per_game = totalpts / totalGames) %>% 
  rename_at("division.x", ~"division") %>% 
  select(-division.y, -totalpts, -totalGames) %>% 
  group_by(division) %>% 
  slice_max(avg_pts_per_game) %>% 
  mutate_if(is.numeric, round)
```

best qb seasons by elo
```{r}
nfl_elo %>% 
  group_by(home_qb, season) %>% 
  summarise(total_home_qb_elo = sum(qb1_value_post)) -> total_home_qb_elo

nfl_elo %>% 
  group_by(away_qb, season) %>% 
  summarise(total_away_qb_elo = sum(qb2_value_post)) -> total_away_qb_elo

totalQBelo <- merge(total_home_qb_elo, total_away_qb_elo, by.x = c("home_qb", "season"), by.y = c("away_qb", "season")) %>% 
  group_by(home_qb, season) %>% 
  summarize(totalQBelo = total_home_qb_elo + total_away_qb_elo) %>% 
  rename_at("home_qb", ~"QB")

totalQBelo

division_df %>% 
  group_by(home_qb, season) %>% 
  summarize(homeQBGames=n()) -> homeQBGames

division_df %>% 
  group_by(away_qb, season) %>% 
  summarize(awayQBGames=n()) -> awayQBGames

totalQBGames <- merge(awayQBGames, homeQBGames, by.x = c("away_qb", "season"), by.y = c("home_qb", "season")) %>% 
  group_by(away_qb, season) %>% 
  summarize(totalQBGames = awayQBGames + homeQBGames) %>% 
  rename_at("away_qb", ~"QB") 

totalQBGames

left_join(totalQBGames, totalQBelo, by = c("QB", "season")) %>% 
  filter(totalQBGames > 8) %>% 
  mutate(avg_qb_elo_per_game = totalQBelo / totalQBGames) %>% 
  select(QB, season, avg_qb_elo_per_game) %>% 
  arrange(-avg_qb_elo_per_game) %>% 
  mutate_if(is.numeric, round)
```

best team seasons by elo
```{r}
nfl_elo %>% 
  group_by(home_team, season) %>% 
  summarise(total_home_elo = sum(elo_home_team)) -> total_home_team_elo

nfl_elo %>% 
  group_by(away_team, season) %>% 
  summarise(total_away_elo = sum(elo_away_team)) -> total_away_team_elo

totalteamelo <- merge(total_home_team_elo, total_away_team_elo, by.x = c("home_team", "season"), by.y = c("away_team", "season")) %>% 
  group_by(home_team, season) %>% 
  summarize(totalelo = total_home_elo + total_away_elo) %>% 
  rename_at("home_team", ~"team")

totalteamelo

division_df %>% 
  group_by(home_team, season) %>% 
  summarize(homeGames=n()) -> homeGames

division_df %>% 
  group_by(away_team, season) %>% 
  summarize(awayGames=n()) -> awayGames

totalGames <- merge(awayGames, homeGames, by.x = c("away_team", "season"), by.y = c("home_team", "season")) %>% 
  group_by(away_team, season) %>% 
  summarize(totalGames = awayGames + homeGames) %>% 
  rename_at("away_team", ~"team") 

totalGames

left_join(totalGames, totalteamelo, by = c("team", "season")) %>%  
  mutate(avg_elo_per_game = totalelo / totalGames) %>% 
  select(team, season, avg_elo_per_game) %>% 
  arrange(-avg_elo_per_game) %>% 
  mutate_if(is.numeric, round)
```
